Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-11 syntaxsemantics.tex
+

After defining the translation in terms of physical load and store action, now we can define our \lstinline|mov| instructions.

%
%\subsection{Machine State under Address Translation}
%\label{sec:selectedinstrsemantics}
%Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation. In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e. address translation traversal as shown in Figure \ref{fig:pagetables}:
%\begin{itemize}
%\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself. The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$. \mytodo{iso: put move}
%\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT). We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
%\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table. Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
%\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
%\end{itemize}

>>>>

After defining the translation in terms of physical load and store action, now we can define our memory \lstinline|mov| instructions
\[
\begin{array}{l}
  \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad \qquad\qquad\qquad 
  \textsf{mov mem64 srcreg64} \\
  \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad \qquad\qquad\qquad 
  \textsf{mov reg64 mem64}
  which are realized by first translating the virtual-memory address to store a value from a register (\TirNameStyle{WriteToMemFromReg}) or load a value at the translated address to be stored in a register (\TirNameStyle{WriteToRegFromMem}). More \lstinline|mov| instructions operating with memory addresses can be found in the submission artifact for different addressing schemes and data sizes.

  Register instructions, on the other hand, are implemented as $\sigma.\mathcal{R}$ lookup and store
\[
\begin{array}{l}
  \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad \qquad\qquad\qquad 
  \textsf{mov ireg64 ireg64} \\
  \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad \qquad\qquad\qquad 
  \textsf{mov creg64 ireg64}
    \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad \qquad\qquad\qquad 
  \textsf{mov ireg64 creg64}
  \end{array}
\]
for different types of registers (e.g. control and general).
%
%\subsection{Machine State under Address Translation}
%\label{sec:selectedinstrsemantics}
%Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation. In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e. address translation traversal as shown in Figure \ref{fig:pagetables}:
%\begin{itemize}
%\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself. The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$. \mytodo{iso: put move}
%\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT). We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
%\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table. Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
%\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
%\end{itemize}
